### 2022-03-27 manuscript.tex
bridgerton [ci skip]
This build is based on
https://github.com/scherzerthesis/thesis/commit/aae47d7eb70a422d565a223f2ab9fc95ccae4700.

This commit was created by the following CI build and job:
https://github.com/scherzerthesis/thesis/commit/aae47d7eb70a422d565a223f2ab9fc95ccae4700/checks
https://github.com/scherzerthesis/thesis/actions/runs/2048527570

This manuscript
(\href{https://scherzerthesis.github.io/thesis/v/0460c57dfce1ec664d567e514d7e4e3dc856b56a/}{permalink})
was automatically generated
from \href{https://github.com/scherzerthesis/thesis/tree/0460c57dfce1ec664d567e514d7e4e3dc856b56a}{scherzerthesis/thesis@0460c57}
on March 24, 2022.

@@

The 5-year survival rate for patients with lung cancer was 25\% in 2020 but varies depending on stage of disease at time of diagnosis (\textbf{doi:10.3322/caac.21654?}).Like other solid cancers, lung cancers are most deadly after primary tumors have metastaized to other organ systems like the liver, kidney, bone, or nervous system (Matsunaga and Shida, 1971). Lung cancer patients with no local or distant metastasis typically have a 5-year survival percentage of approximately 40\% compared to approximately 5\% with distant metastasis{[}Woodard \emph{et al.}; 27740970{]}. Ultimately, the cause of death of many lung cancer patients include primary lung tumor burden, infection, pulmonary embolism or hemorrhage, or complications of metastatic dissemination(Nichols \emph{et al.}, 2012). Although patient survival has improved due to improved targeted and immune therapies, the still poor prognosis reflects a need to better understand the molecular mechanisms underlying lung adenocarcinoma initiation, progression, maintenance and response to targeted- or immune-therapy.

@@

To determine the size of BRAF \textsuperscript{V600E} tumors and the contribution of 11 tumor suppressor genes (\emph{Trp53, Rb1, Cdkn2a, Apc, Rbm10, Stk11, Setd2, Arid1a, Atm, Keap1, Smad4)} on BRAF \textsuperscript{V600E} lung tumorigenesis, we obtained tittered barcoded-lentiviral plasmids that encode Cre-recombinase as well as a sgRNA that inactivates one of eleven tumor suppressors common in lung adenocarcinoma. All genes tested are putative tumor suppressor genes commonly lost in human lung cancer patients. Each lentiviral solution contained \textasciitilde1000 individually barcoded lentiviruses. We infected both \emph{LSL-Cas9} and \emph{CAS9} deficient mice in order to determine the contribution of genetic manipulation and/or deletion of the targeted gene. Mice meant to be analyzed after 8 weeks received a viral titer of 1x10\^{}5 while mice mean to be analyzed after 16 weeks received a viral titer of 1x10\^{}4. After euthanasia, lungs were collected for genomic DNA extraction, and barcode amplification followed by preperation of sequencing libraries. Before genomic extraction and library preparation, a known quantity of known-barcoded cells was added in order to benchmark sequencing reads to a known cell number.

@@

@@

@@

all cell types in the lung
tumor cell clusters
highlight known cell types
repspone to D+T each cluster?

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-8Gf0ScB2}{}}%
(2014) \href{https://doi.org/10.1038/nature13385}{Comprehensive molecular profiling of lung adenocarcinoma.} \emph{Nature}, \textbf{511}, 543--50.

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

This manuscript
(\href{https://scherzerthesis.github.io/thesis/v/aae47d7eb70a422d565a223f2ab9fc95ccae4700/}{permalink})
was automatically generated
from \href{https://github.com/scherzerthesis/thesis/tree/aae47d7eb70a422d565a223f2ab9fc95ccae4700}{scherzerthesis/thesis@aae47d7}
on March 27, 2022.

@@

The 5-year survival rate for patients with lung cancer was 25\% in 2020 but varies depending on stage of disease at time of diagnosis (Siegel \emph{et al.}, 2021).Like other solid cancers, lung cancers are most deadly after primary tumors have metastaized to other organ systems like the liver, kidney, bone, or nervous system (Matsunaga and Shida, 1971). Lung cancer patients with no local or distant metastasis typically have a 5-year survival percentage of approximately 40\% compared to approximately 5\% with distant metastasis{[}Woodard \emph{et al.}; 27740970{]}. Ultimately, the cause of death of many lung cancer patients include primary lung tumor burden, infection, pulmonary embolism or hemorrhage, or complications of metastatic dissemination(Nichols \emph{et al.}, 2012). Although patient survival has improved due to improved targeted and immune therapies, the still poor prognosis reflects a need to better understand the molecular mechanisms underlying lung adenocarcinoma initiation, progression, maintenance and response to targeted- or immune-therapy.

@@

Adenoviral vectors and lentiviral vectors introduced in the lungs of GEMMs offer temporal control and can be titrated to acheive an optimal level of tumor numerosity. However, only lentiviral vectors can be used to barcode tumors because they integrate into the genome while adenoviruses do not(DuPage \emph{et al.}, 2009; Rogers \emph{et al.}, 2017). Together with CRISPR/CAS9 gene editing, lentiviruses have been used to analyzed many genetic interactions of oncogenic \emph{Kras} driven lung adenocarcinoma(Winters \emph{et al.}, 2017; Rogers \emph{et al.}, 2017). To enable simultaneous actvation of BRAF\textsuperscript{V600E} and loss of tumor supressor genes, we bred \emph{Braf\textsuperscript{CAT}}; \emph{H11b\textsuperscript{LSL-CAS9/+}} mice.
To determine the size of BRAF \textsuperscript{V600E} tumors and the contribution of 11 tumor suppressor genes (\emph{Trp53, Rb1, Cdkn2a, Apc, Rbm10, Stk11, Setd2, Arid1a, Atm, Keap1, Smad4)} on BRAF \textsuperscript{V600E} lung tumorigenesis, we obtained tittered barcoded-lentiviral plasmids that encode Cre-recombinase as well as a sgRNA that inactivates one of eleven tumor suppressors common in lung adenocarcinoma. All genes tested are putative tumor suppressor genes commonly lost in human lung cancer patients. Each lentiviral solution contained \textasciitilde1000 individually barcoded lentiviruses. We infected both \emph{LSL-Cas9} and \emph{CAS9} deficient mice in order to determine the contribution of genetic manipulation and/or deletion of the targeted gene. Mice meant to be analyzed after 8 weeks received a viral titer of 1x10\^{}5 while mice mean to be analyzed after 16 weeks received a viral titer of 1x10\^{}4. After euthanasia, lungs were collected for genomic DNA extraction, and barcode amplification followed by preperation of sequencing libraries. Before genomic extraction and library preparation, a known quantity of known-barcoded cells was added in order to benchmark sequencing reads to a known cell number.

@@

Our Tuba-Seq analysis revealed that there is a spectrum of BRAF\textsuperscript{V600E} tumors after 16-weeks. This observation would support the idea that just because a cell is activated by BRAF\textsuperscript{V600E}, does not mean

@@

The mammalian lung is a vast and intricate labyrinth that is responsible that is responsible for gas exchange between air and body through a meshwork of endothelial cells. The lung organ system begins with a single airway that branches in a geometric and stereotypic pattern(Metzger \emph{et al.}, 2008). These many thousands of branches lead to the business end of gas exchange, the lung alveoli. The alveoli are comprised of two specialized cell types: Alveolar Type 1 (AT1) cells comprise most of the surface area of the lung and are thin squamous cells that mediate oxygen exchange across cell membranes. Alveolar Type 2 (AT2) cells, however, are cuboidal that secrete surfactant proteins that aide in preventing the delicate alveoli from collapsing.

Pathology of the alveoli is a steadfast threat to an organism's life due to the importance of gas exchange. Diseases such as lung adenocarcinoma, pulmonary fibrosis, and emphysema are the most common dysfunctions of the distal lung epithelium(Desai \emph{et al.}, 2014). Lung adenocarcinoma is also the most common type of human lung cancers in both men and women. A large body of literature has suggested that lung adenocarcinomas arise in the alveolar and are descendants of AT2 cells (Ferone \emph{et al.}, 2020; Jackson \emph{et al.}, 2001). Therefore, there is great interest in understanding how AT2 cells develop, maintain cell identity, as well as how they renew themselves during aging or in response to injury (Treutlein \emph{et al.}, 2014; Alanis \emph{et al.}, 2014, 2014; 2020).

Recent advances in profiling of single-cell transcriptomes during homeostasis conditions and diseases states have greatly improved our understanding of AT2 cell, and lung adenocarcinoma cell biology(Marjanovic \emph{et al.}, 2020; LaFave \emph{et al.}, 2020; Treutlein \emph{et al.}, 2014).Single-cell profiling of human lung cancers have revealed the drastic intra-tumoral heterogeneity that exists in a single tumor(Gerlinger \emph{et al.}, 2012; Xue \emph{et al.}, 2016). Since transcriptional heterogeneity can be considered a source of therapeutic resistance, there is great interest in understanding the evolution of lung adenocarcinoma and how it changes with response to pathway targeted therapy(Maynard \emph{et al.}, 2020).Since GEMMs are an important pre-clinical model for testing out potential therapeutic strategies for lung adenocarcinoma, it is important to understand single-cell heterogeneity in popular GEMM models.

Although GEMMs are a great way to model human lung adenocarcinomas, there are several drawbacks worth mentioning. Human lung adenocarcinomas typically develop from a single tumor-initiating cell and is a consequence of an accumulation of mutations in certain cell types and a weakened immune system across a lifetime. In GEMMs, tumorigenesis is often initiated in many cells at once and is the consequence of an engineered mutation. In contrast to human lung adenocarcinomas that have a high mutational burden, murine lung adenocarcinomas do not have a high mutational burden and instead display high copy number variation(Westcott \emph{et al.}, 2014). Therefore, we set out to profile the single-cell transcriptomes of various BRAF\textsuperscript{V600E}-driven lung cancer GEMMs in order to determine genotype-specific heterogeneity and response to pathway-targeted therapy.

@@

Adenoviral vectors and lentiviral vectors introduced in the lungs of GEMMs offer temporal control and can be titrated to acheive an optimal level of tumor numerosity. However, only lentiviral vectors can be used to barcode tumors because they integrate into the genome while adenoviruses do not(DuPage \emph{et al.}, 2009; Rogers \emph{et al.}, 2017).

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-179MWicwA}{}}%
(2020) \href{https://doi.org/10.1038/s41586-020-2496-1}{A single-cell transcriptomic atlas characterizes ageing tissues in the mouse.} \emph{Nature}, \textbf{583}, 590--595.

\leavevmode\vadjust pre{\hypertarget{ref-8Gf0ScB2}{}}%
(2014) \href{https://doi.org/10.1038/nature13385}{Comprehensive molecular profiling of lung adenocarcinoma.} \emph{Nature}, \textbf{511}, 543--50.

@@

\leavevmode\vadjust pre{\hypertarget{ref-vsbCfYx4}{}}%
Alanis,D.M. \emph{et al.} (2014) \href{https://doi.org/10.1038/ncomms4923}{Two nested developmental waves demarcate a compartment boundary in the mouse lung.} \emph{Nat Commun}, \textbf{5}, 3923.

@@

\leavevmode\vadjust pre{\hypertarget{ref-n08HchE9}{}}%
Desai,T.J. \emph{et al.} (2014) \href{https://doi.org/10.1038/nature12930}{Alveolar progenitor and stem cells in lung development, renewal and cancer.} \emph{Nature}, \textbf{507}, 190--4.

@@

\leavevmode\vadjust pre{\hypertarget{ref-bhNe3oCb}{}}%
DuPage,M. \emph{et al.} (2009) \href{https://doi.org/10.1038/nprot.2009.95}{Conditional mouse lung cancer models using adenoviral or lentiviral delivery of Cre recombinase.} \emph{Nat Protoc}, \textbf{4}, 1064--72.

@@

\leavevmode\vadjust pre{\hypertarget{ref-1EmZp72ZM}{}}%
Gerlinger,M. \emph{et al.} (2012) \href{https://doi.org/10.1056/nejmoa1113205}{Intratumor heterogeneity and branched evolution revealed by multiregion sequencing.} \emph{N Engl J Med}, \textbf{366}, 883--892.

@@

\leavevmode\vadjust pre{\hypertarget{ref-RCmUro9f}{}}%
LaFave,L.M. \emph{et al.} (2020) \href{https://doi.org/10.1016/j.ccell.2020.06.006}{Epigenomic State Transitions Characterize Tumor Progression in Mouse Lung Adenocarcinoma.} \emph{Cancer Cell}, \textbf{38}, 212--228.e13.

@@

\leavevmode\vadjust pre{\hypertarget{ref-btzNHt9A}{}}%
Maynard,A. \emph{et al.} (2020) \href{https://doi.org/10.1016/j.cell.2020.07.017}{Therapy-Induced Evolution of Human Lung Cancer Revealed by Single-Cell RNA Sequencing.} \emph{Cell}, \textbf{182}, 1232--1251.e22.

@@

\leavevmode\vadjust pre{\hypertarget{ref-OuXi55DW}{}}%
Metzger,R.J. \emph{et al.} (2008) \href{https://doi.org/10.1038/nature07005}{The branching programme of mouse lung development.} \emph{Nature}, \textbf{453}, 745--50.

@@

\leavevmode\vadjust pre{\hypertarget{ref-gY2NonCA}{}}%
Siegel,R.L. \emph{et al.} (2021) \href{https://doi.org/10.3322/caac.21654}{Cancer Statistics, 2021}. \emph{CA A Cancer J. Clin.}, \textbf{71}, 7--33.

@@

\leavevmode\vadjust pre{\hypertarget{ref-NLPwvKCP}{}}%
Treutlein,B. \emph{et al.} (2014) \href{https://doi.org/10.1038/nature13173}{Reconstructing lineage hierarchies of the distal lung epithelium using single-cell RNA-seq.} \emph{Nature}, \textbf{509}, 371--5.

@@

\leavevmode\vadjust pre{\hypertarget{ref-2gOIsLdX}{}}%
Winters,I.P. \emph{et al.} (2017) \href{https://doi.org/10.1038/s41467-017-01519-y}{Multiplexed in vivo homology-directed repair and tumor barcoding enables parallel quantification of Kras variant oncogenicity.} \emph{Nat Commun}, \textbf{8}, 2053.

@@

\leavevmode\vadjust pre{\hypertarget{ref-FfxdQ0tV}{}}%
Xue,Y. \emph{et al.} (2016) \href{https://doi.org/10.1007/s13238-016-0330-1}{Evolution from genetics to phenotype: reinterpretation of NSCLC plasticity, heterogeneity, and drug resistance.} \emph{Protein Cell}, \textbf{8}, 178--190.
